---
title: "Peer Assessment 1"
output:
  html_document: default
  pdf_document: default
---

loading the dataset 
```{r setup, include=FALSE}
setwd("C://Users//pg000//Desktop//Download//")

data1 <- read.csv("activity.csv")

library(ggplot2)
library(knitr)
library(dplyr)

head(data1,10)

```

## the number of step according to the date
## make histogram that the numbers steps taken each day 
```{r }
step_date <- aggregate(steps~date, data = data1,FUN = sum, na.rm=FALSE)
str(step_date)

ggplot(step_date, aes(x=steps))+
 geom_histogram(bins = 20)+
 labs(title = "total number of steps", 
      x = "Number of steps of per day",
      y = "Number of times in day")
```

## calculate the mean and median number 

```{r pressure, echo=FALSE}
mean_step <- mean(step_date$steps)
mean_step
summary(step_date)
```

## intervals that average of number of steps taken 
## plot in time series 
```{r}
interval_step <- aggregate(steps ~ interval, data1, mean)
head(interval_step)

ggplot(interval_step, aes(x =interval , y =steps))+
 geom_line()+
 labs(title = "average daily activity pattern",
      x ="interval", y= "steps")
```
## finding max number
```{r}
max_steps <- which.max(interval_step$steps)

interval_step[max_steps,]
```
## imputing missing value 
```{r}
sum(is.na(data1))

data_filled <- data1 
means <- mean(data1$steps, na.rm  =TRUE)
data_filled$steps[is.na(data_filled$steps)] <- means

str(data_filled)
```
## plot the dataset when missing value filled in 

```{r}
steps_datafill <- aggregate(steps~date, data = data_filled, FUN = sum)
head(steps_datafill)

ggplot(steps_datafill, aes(x=steps))+
 geom_histogram(bins = 20)+
 labs(title = "Histogram of the total number of steps",
      x = "Number of steps per day", y ="Number times in day")

summary(step_date)
summary(steps_datafill)
```
## check differences in activity patterns between weekdays and weekends and plot it
```{r }
dataNew <- data_filled
dataNew["type_of_day"] <- weekdays(as.Date(dataNew$date))
dataNew$type_of_day[dataNew$type_of_day %in% c('Saturday','Sunday')] <- 'weekend'
dataNew$type_of_day[dataNew$type_of_day != 'weekend'] <- 'weekday'

dataNew$type_of_day <- as.factor(dataNew$type_of_day)
dataNew_SI <- aggregate(steps~interval + type_of_day, dataNew, mean)
head(dataNew_SI)   

ggplot(dataNew_SI, aes(x = interval , y =steps))+
 geom_line()+
 labs(title = "Average daily steps",
      x = "interval",
      y = "Total number of steps")+
 facet_wrap(~type_of_day,ncol = 1, nrow =2)
```